---
title: "Activity Analysis"
author: Rajib Kar
date: March 25, 2017
output: md_document
---


Analysing Activity monitoring data
=============================================
Author:-Rajib Kar
-----------------

Introduction
-------------
It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.


Data
-----
The variables included in this dataset are:

Steps: Number of steps taking in a 5-minute interval.

Date: The date on which the measurement was taken in YYYY-MM-DD format.

Interval: Identifier for the 5-minute interval in which measurement was taken.

Questions
---------
1.What is mean total number of steps taken per day?

2.What is the average daily activity pattern?

3.Are there differences in activity patterns between weekdays and weekends?

Loading the data
--------------------
```{r Download}
download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",destfile = "Steps.zip")
unzip("Steps.zip")
act<-read.csv("activity.csv")
```
Calculate the total number of steps taken per day.
-------------------------------------------------
```{r}
library(dplyr)

act$date<-as.Date(as.character(act$date),"%Y-%m-%d")

act_total_steps<-act%>%group_by(date)%>%summarise(total_steps=sum(steps,na.rm=TRUE))

print(act_total_steps)
```
Making Histogram
--------------------
```{r}
options(warn=-1)
library(ggplot2)
ggplot(act_total_steps,aes(x=date,y=total_steps))+geom_histogram(stat = "identity",fill="blue")+ggtitle("Total number of steps taken per day") + 
  xlab("Date") + 
  ylab("Steps")
```


The mean and median of the total number of steps taken each day.
----------------------------------------------------------------

```{r}
library(dplyr)

act_total_steps<-act%>%group_by(date)%>%summarise(total_steps=sum(steps,na.rm=TRUE))

mean(act_total_steps$total_steps)

median(act_total_steps$total_steps)

```

Plotting The activity Pattern Per Day
-------------------------------------

```{r}
act_avg_steps<-act%>%group_by(interval)%>%summarise(avg_steps=round(mean(steps,na.rm=TRUE)))

with(act_avg_steps,plot(x=c(1:288),y=avg_steps,type = "l",col="red",xlab = "Time interval",ylab = "Average steps",main ="Average Steps per Interval" ))
```

Imputing missing values
-----------------------
total number of missing values
```{r}
table(is.na(act$steps))
```

All of the missing values are filled in with mean value for that 5-minute interval.
----------------------------------------------------------------------------------
```{r}
act_new<-act

for(i in 1:nrow(act_new))
{
 if(is.na(act_new[i,]$steps))
 {
   int<-act_new[i,]$interval
   act_new[i,]$steps<-subset(act_avg_steps,interval==int)$avg_steps
 }
}
table(is.na(act_new$steps))
```

Making Histogram after handLing the missing values
-------------------------------------------------
```{r}
options(warn=-1)
library(ggplot2)
library(dplyr)

act_total_steps_new<-act_new%>%group_by(date)%>%summarise(total_steps=sum(steps,na.rm=TRUE))

ggplot(act_total_steps_new,aes(x=date,y=total_steps))+geom_histogram(stat = "identity",fill="blue")+ggtitle("Total number of steps taken per day") + 
  xlab("Date") + 
  ylab("Steps")
```


Recalculating  mean and median of the total number of steps taken each day.
----------------------------------------------------------------

```{r}


mean(act_total_steps_new$total_steps)

median(act_total_steps_new$total_steps)

```

Calculating activity patterns in Weekdays and Weekends
------------------------------------------------------

```{r}
act_new$Day_Type<-"Weekday"
act_new[weekdays(act_new$date)%in% c("Sunday","Saturday"),]$Day_Type<-"Weekend"
act_new$Day_Type<-as.factor(act_new$Day_Type)
act_avg_steps_new<-act_new%>%group_by(interval,Day_Type)%>%summarize(avg_steps=mean(steps))

library(ggplot2)
g<-ggplot(act_avg_steps_new,aes(x=interval,y=avg_steps))

g+geom_line(aes(col=Day_Type))+facet_grid(.~Day_Type)+labs(xlab("Interval"))+labs(ylab("Average Steps"))
```
